An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-03-24 main.tex
Updates from Overleaf

%%%M
\subsection*{Details of left to right direction of Theorem~\ref{theo1}.}
One can translate any QLTL property into a second-order monadic logic in a standard
way~\cite{Thomas}, and then from
second-order monadic logic to a deterministic Muller automaton $M = (S, \iota , \Sigma, \Delta, {\cal F})$ (a B\"{u}chi automaton may not always be deterministic), where $S$ is the set of states, $\iota \in S$ is the initial state, $\Sigma$ is the set of inputs, each $\alpha \in \Sigma$ is a minterm (i.e., conjunction of
each variable in the set negated or non-negated) of the variables in $B$, $\Delta : S \times \Sigma \mapsto S$ is the transition function, and ${\cal F} = \{ F_1 , \ldots , F_k \}$ is the accepting set. A muller automaton 
accepts a run if it 
passes infinitely often in exactly
one of the specified accepting sets. To formulate an equivalent ELTL formula,
the auxiliary variables represent the states $S$ of $M$, i.e., $a_s$ represents the state $s \in S$.

%\klaus{The following formulation above is not not clear: ``where an accepted input needs to pass infinitely often in exactly
%one of the specified sets''.}

\noindent
Then we translate $M$ into an ELTL formula
of the form:

@@

\begin{itemize}
\item $I [ \varphi , {\cal R} , 0 ] = \emptyset$.
\item $I [ \true , {\cal R} , i ] = \{ \epsilon \}$.
\item $I [ p ( a ) , {\cal R} , i ] =$ if $m (p )  ( a , i ) $ then
$\{ \epsilon \}$ else $\emptyset$.
\item $I [ p ( x ) , {\cal R} , i ] = \{ [ x \mapsto a ] | m (p) ( a , i)  \}$.
%\item $I [ \true , \sigma , i ] = D^n$.
%\item $I [ ( \varphi \vee \psi ) , \sigma , i ] = 
%I [ \varphi , \sigma , i ] \cup I [ \psi , \sigma , i ]$.
\item $I [ ( \varphi \wedge \psi ) , \sigma , i ] = 
I [ \varphi , {\cal R} , i ] \;  \bigcap \; I [ \psi , \sigma , i ]$.
\item $I [ \neg \varphi , {\cal R} , i ] = 
A_{\vars ( \varphi )} \; \setminus \; I [ \varphi , {\cal R} , i ]$.
\item $I [ ( \varphi \; \since \; \psi ) , {\cal R} , i ] = 
I [ \psi , {\cal R}, i ] \; \bigcup \;
( I [ \varphi , {\cal R} , i ] \bigcap 
I [ (\varphi \since \psi ) , {\cal R} , i - 1 ] )$.
\item $I [ \ominus \varphi , {\cal R} , i ] = I [ \varphi , {\cal R} , i-1 ]$.
\item $I [ \exists x \; \varphi , \sigma , i ] = 
\proj ( I [ \varphi , {\cal R} , i ] , \{ x \} )$.
\item $I [ \psi \mathrm{\ where\ } r_j ( x_j) := 
\varphi_j (x_j) : j \in \{ 1 , \ldots , n \} , {\cal R} , i ] = 
 I [ \psi , {\cal Q} ] $ where ${\cal Q} = {\cal R} \cup
\{ r_j | r_j [ i ] =  I [ 
\varphi_j , {\cal Q} , i ] \}$
\end{itemize}
Note that the last time interprets a formula of
the form defined in~(\ref{EFLTL}).

%\klaus{Above: as the semantics is presented, it looks as if 
%the `$\psi\ where ...$' is just another subformula.}

%\klaus{Above: hide is not defined.}

%\klaus{Above: the semantics of $p(a)$ and $p(x)$ do not refer to 
%${\cal R} $.}

In the last item, the extended model ${\cal Q}$ appears both on the left and the right of the equality sign. It is well defined because of Lemma~\ref{samesame}.
\end{document}

>>>>

%%%M
\subsection*{Details of left to right direction of Theorem~\ref{theo1}.}
One can translate any QLTL property into a second-order monadic logic in a standard
way~\cite{Thomas}, and then from
second-order monadic logic to a deterministic Muller automaton $M = (S, \iota , \Sigma, \Delta, {\cal F})$ (a B\"{u}chi automaton may not always be deterministic), where $S$ is the set of states, $\iota \in S$ is the initial state, $\Sigma$ is the set of inputs, each $\alpha \in \Sigma$ is a minterm (i.e., conjunction of
each variable in the set negated or non-negated) of the variables in $B$, $\Delta : S \times \Sigma \mapsto S$ is the transition function, and ${\cal F} = \{ F_1 , \ldots , F_k \}$ is the accepting set. A muller automaton 
accepts a run if it 
passes infinitely often in exactly
one of the specified accepting sets. To formulate an equivalent ELTL formula,
the auxiliary variables represent the states $S$ of $M$, i.e., $a_s$ represents the state $s \in S$.
Then we translate $M$ into an ELTL formula
of the form:

@@

\begin{itemize}
\item $I [ \varphi , {\cal R} , 0 ] = \emptyset$.
\item $I [ \true , {\cal R} , i ] = \{ \epsilon \}$.
\item $I [ p ( a ) , {\cal R} , i ] =$ if $m (p )  ( a , i ) $ then
$\{ \epsilon \}$ else $\emptyset$.
\item $I [ p ( x ) , {\cal R} , i ] = \{ [ x \mapsto a ] | m (p) ( a , i)  \}$.
%\item $I [ \true , \sigma , i ] = D^n$.
%\item $I [ ( \varphi \vee \psi ) , \sigma , i ] = 
%I [ \varphi , \sigma , i ] \cup I [ \psi , \sigma , i ]$.
\item $I [ ( \varphi \wedge \psi ) , \sigma , i ] = 
I [ \varphi , {\cal R} , i ] \;  \bigcap \; I [ \psi , \sigma , i ]$.
\item $I [ \neg \varphi , {\cal R} , i ] = 
A_{\vars ( \varphi )} \; \setminus \; I [ \varphi , {\cal R} , i ]$.
\item $I [ ( \varphi \; \since \; \psi ) , {\cal R} , i ] = 
I [ \psi , {\cal R}, i ] \; \bigcup \;
( I [ \varphi , {\cal R} , i ] \bigcap 
I [ (\varphi \since \psi ) , {\cal R} , i - 1 ] )$.
\item $I [ \ominus \varphi , {\cal R} , i ] = I [ \varphi , {\cal R} , i-1 ]$.
\item $I [ \exists x \; \varphi , \sigma , i ] = 
\proj ( I [ \varphi , {\cal R} , i ] , \{ x \} )$.
\item $I [ \psi \mathrm{\ where\ } r_j ( x_j) := 
\varphi_j (x_j) : j \in \{ 1 , \ldots , n \} , {\cal R} , i ] = 
 I [ \psi , {\cal Q} ] $ where ${\cal Q} = {\cal R} \cup
\{ r_j | r_j [ i ] =  I [ 
\varphi_j , {\cal Q} , i ] \}$
\end{itemize}
Note that the last time interprets a formula of
the form defined in~(\ref{EFLTL}).
%\klaus{Above: as the semantics is presented, it looks as if 
%the `$\psi\ where ...$' is just another subformula.}
%\klaus{Above: hide is not defined.}
%\klaus{Above: the semantics of $p(a)$ and $p(x)$ do not refer to 
%${\cal R} $.}
In the last item, the extended model ${\cal Q}$ appears both on the left and the right of the equality sign. It is well defined because of Lemma~\ref{samesame} (equivalently, a more verbose definition can include a stepwise construction of ${\cal Q}$ from ${\cal R}$).
\end{document}
